Missing The Point

### 2020-10-22 Manuscript.Rmd
Re-run sim with new R version, fix non-convergence in case study

```{r setup, include=FALSE}
# environment
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mice)
load("Data/results_25_75.Rdata")
# load data
load("Data/example.Rdata")
load("Data/example_mids.Rdata")

# sim results plotting function
simulation_plot <- function(v, name, ...){
  results %>%
    mutate(
      p = factor(p, levels = c(0.25, 0.5, 0.75), labels = c("25%", "50%", "75%")),
      mech = factor(mech, levels = c("MCAR", "MAR", "MNAR")),
      est = est - 1,
      rsq = rsq - 0.5
      ) %>% 
  ggplot(aes(
    x=.data[["it"]], 
    y=.data[[v]], 
    color = .data[["p"]], 
    shape = .data[["mech"]], 
    linetype = .data[["mech"]]
    )) +
  geom_line(na.rm = TRUE) +
  labs(
    x = "Number of iterations",
    y = name,
    color = "Incomplete \n cases", 
    shape = "Missingness \n mechanism", 
    linetype = "Missingness \n mechanism") +
  theme_classic() + 
  scale_color_manual(values = c(
    #"#009E73", #green
    "#56B4E9", #blue
    #"#E69F00", #orange
    "#F0E442", #yellow
    "#D55E00"  #red
  ))
  # scale_color_manual(values = c(
  #   # '#228833', #green
  #   # '#66CCEE', #blue
  #   '#CCBB44', #yellow
  #   '#EE6677',  #red~ish
  #   '#AA3377'  #maroon
  #   ))
}

# plot each var of example data
casestudy_plot <- function(v, name, ...){
ggplot(example, aes(x = .data[["it"]], y = .data[[v]])) +
  geom_line(na.rm = TRUE) +
    labs(x = "Iteration number",
         y = name) +
  theme_classic()
}

@@

With iterative imputation, the validity of the inference depends on the state-space of the algorithm at the final iteration. This introduces a potential threat to the validity of the imputations: What if the algorithm has not converged? Are the imputations then to be trusted? And can we rely on the inference obtained using the imputed data? These remain open questions since the convergence properties of iterative imputation algorithms have not been systematically studied (Van Buuren, 2018, 6.5.2). While there is no scientific consensus on how to evaluate the convergence of imputation algorithms [@zhu15; @taka17], the current practice is to visually inspect imputations for signs of non-convergence. 
<!-- (Zhu & Raghunathan, 2015; Takahashi, 2017) -->

@@

-   What does non-convergence look like? --\> non-stationarity + non-mixing in a certain parameter, e.g., chain means.

-   What are the consequences? --\> bias, under-coverage --\> refer to van Buuren (2018) instead of reproducing this

@@

The following figures provide an overview of the results for each method and each performance measure, contrasted to the number of iterations in the imputation algorithm. Within the figures, we split the results according to the missingness conditions [missingness mechanisms as line types, and proportion of incomplete cases as colors].

@@

In this figure, we see autocorrelation in the chain means as the diagnostic method. We can interpret this as the extent to which the average imputed value trends across iterations [unnecessary??]. We observe that autocorrelation in the chain means rapidly decreases until $it \geq 6$. This means that there is some initial trending within chains, but stationarity does not improve substantively after the first few iterations. These results hold irrespective of the missingness condition.

@@

```{r echo=FALSE}
p <- plot(mids)
p[1:2,1] # %>% plotly::ggplotly()
```

@@

This figure shows that 36 iterations are required before the $\widehat{R}$-values of the chain means drop below the threshold for non-convergence.

@@

The scientific estimate reaches the threshold much sooner, when $it=15$.

@@

According to the $\widehat{R}$-values of the novel parameter, at least 25 iterations are required.

>>>>

```{r setup, include=FALSE}
# environment
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mice)
load("Data/results_25_75.Rdata")
# load data
load("Data/example.Rdata")
load("Data/example_chains.Rdata")

# sim results plotting function
simulation_plot <- function(d = results, v, name, ...){
  d %>%
    mutate(
      p = factor(p, levels = c(0.25, 0.5, 0.75), labels = c("25%", "50%", "75%")),
      mech = factor(mech, levels = c("MCAR", "MAR", "MNAR")),
      est = est - 1,
      rsq = rsq - 0.5
      ) %>% 
  ggplot(aes(
    x=.data[["it"]], 
    y=.data[[v]], 
    color = .data[["p"]], 
    shape = .data[["mech"]], 
    linetype = .data[["mech"]]
    )) +
  geom_line(na.rm = TRUE) +
  labs(
    x = "Number of iterations",
    y = name,
    color = "Incomplete \n cases", 
    shape = "Missingness \n mechanism", 
    linetype = "Missingness \n mechanism") +
  theme_classic() + 
  scale_color_manual(values = c(
    #"#009E73", #green
    "#56B4E9", #blue
    #"#E69F00", #orange
    "#F0E442", #yellow
    "#D55E00"  #red
  ))
  # scale_color_manual(values = c(
  #   # '#228833', #green
  #   # '#66CCEE', #blue
  #   '#CCBB44', #yellow
  #   '#EE6677',  #red~ish
  #   '#AA3377'  #maroon
  #   ))
}

# plot each var of example data
casestudy_plot <- function(d = example, v, name, ...){
ggplot(d, aes(x = .data[["it"]], y = .data[[v]])) +
  geom_line(na.rm = TRUE) +
    labs(x = "Iteration number",
         y = name) +
  theme_classic()
}

@@

With iterative imputation, the validity of the inference depends on the state-space of the algorithm at the final iteration. This introduces a potential threat to the validity of the imputations: What if the algorithm has not converged? Are the imputations then to be trusted? And can we rely on the inference obtained using the imputed data? These remain open questions since the convergence properties of iterative imputation algorithms have not been systematically studied (Van Buuren, 2018, 6.5.2). While there is no scientific consensus on how to evaluate the convergence of imputation algorithms [@liu14; @zhu15; @taka17], the current practice is to visually inspect imputations for signs of non-convergence. 
<!-- (Zhu & Raghunathan, 2015; Takahashi, 2017) -->

@@

- With MCMC, the generated values will vary even after convergence, so there is not a unique point at which convergence is established [@gelm13]. Therefore, diagnostic methods may only identify signs of *non*-convergence [@hoff09]. Non-convergence occurs when one of two requirements for convergence is not met.

- There are two requirements for convergence of iterative algorithms: mixing and
stationarity [@gelm13]. Mixing implies that generated values intermingle nicely, and stationarity is characterized by the absence of trending between successive draws.

-   What are the consequences? --\> bias, under-coverage [refer to van Buuren (2018) instead of reproducing this]

@@

The following figures display the simulation results for each method and each performance measure, contrasted to the number of iterations in the imputation algorithm. Within the figures, we split the results according to the missingness conditions [missingness mechanisms as line types, and proportion of incomplete cases as colors].

@@

We observe that autocorrelation in the chain means rapidly decreases until $it \geq 6$. This means that there is some initial trending within chains, but stationarity of the average imputed value does not improve substantively after the first few iterations. These results hold irrespective of the missingness condition.

@@

```{r echo=FALSE}
# p <- plot(mids)
# p[1:2,1] # %>% plotly::ggplotly()
param %>% 
  ggplot() + 
  geom_line(aes(x = it, y = mu.hgt, color = as.factor(m))) + 
  theme_classic() + 
  labs(x = "Iteration number",
       y = "Chain mean variable 'hgt'",
       color = "Imputation")
```

@@

This figure shows that 7 iterations are required before the $\widehat{R}$-values of the chain means drop below the threshold for non-convergence.

@@

The scientific estimate reaches the threshold much sooner, when $it=14$.

@@

According to the $\widehat{R}$-values of the novel parameter, at least 15 iterations are required.
